### 2023-03-28 definitions.tex
details

\section{Definitions}\label{section definitions}

@@

\begin{definition}
An {\em transition system} is a pair $\mathscr{S} = (S,\rightarrow )$ where $S$ is a set of 
{\em states} and  
$ {\rightarrow} \subseteq S \times S$ is a
binary relation 
on
the set of states, denoted as the set of {\em transitions}. A {\em labeled transition system} (LTS for short) is a tuple $\mathscr{S}=(S, \Sigma, \rightarrow)$ where 
$S$ is a set of {\em states}, $\Sigma$ is a (finite) set of {\em labels}, and 
${\rightarrow} \subseteq S\times \Sigma \times S$ is a 
ternary relation,
denoted as the set of {\em labeled transitions}. 
\end{definition}

Note 
 that an (unlabeled) transition system can be seen as a labeled transition system where the set of labels consists of only one element.

We again prefer to use infix notation and write $s \rightarrow s'$ to denote a {\em transition} from state $s$ to state $s'$ (i.e., $ (s,s') \in  {\rightarrow} $). \newline

%
\alain{enlever ce qui est inutile}
%
Determinism, (infinite) paths, their length, and concatenation in unlabeled transition systems are
then defined as expected.

@@

We write $\rightarrow^{k}$, $\rightarrow^{+}$, $\rightarrow^{=}$, $\rightarrow^{*}$
for the $k$-step iteration of $\rightarrow$, its transitive closure, its reflexive closure, its reflexive and transitive closure). We use similar notation for $\post$ and $\pred$...

Let $X,Y \subseteq S$ and $k \in \mathbb{N}$. We denote $X \longrightarrow^{*} Y$ (resp. $X \longrightarrow^{\leq k} Y$) if from all states $x \in X$ there exists a path (resp. of length smaller than $k$) that reaches a state $y \in Y$.

@@

\noindent
The set of {\em (immediate) successors} of a state $s \in S$ is deﬁned as
$\post(s) = \{ s' \in S \mid \exists a \in \Sigma ~ s \xrightarrow{a} s'\}$.
A state without successors is called a {\em dead end}.  \\
The set of {\em (immediate) predecessor} of a state $s \in S$ is deﬁned as
$\pred(s) = \{ s' \in S \mid \exists a \in \Sigma ~ s' \xrightarrow{a} s\}$.

@@

We abbreviate $x \leq y \not\leq x$ by $x < y$.

% A partial ordering (a po) is an antisymmetric qo

Any qo induces an equivalence relation ($x \equiv y$ iff $x \leq y \leq x$).
% and gives rise to a po between the equivalence classes ?
% 	do we need po ?

@@

\begin{definition}
 A {\em well-quasi-ordering} (a wqo) is any quasi-ordering $(X,\leq)$ such that, for any infinite sequence $x_0, x_1, x_2, ...$ in $X$, there exist indexes $i \leq j$ with
$x_i \leq  x_j$.
\end{definition}

>>>>

\section{Well-structured transition systems}\label{section definitions}

@@

\begin{definition}
An {\em transition system} is a pair $\mathscr{S} = (S,\rightarrow )$ where $S$ is a set of 
{\em states} and  
$ {\rightarrow} \subseteq S \times S$ is a
binary relation 
on
the set of states, denoted as the set of {\em transitions}.

%A {\em labeled transition system} (LTS for short) is a tuple $\mathscr{S}=(S, \Sigma, \rightarrow)$ where 
%$S$ is a set of {\em states}, $\Sigma$ is a (finite) set of {\em labels}, and 
%${\rightarrow} \subseteq S\times \Sigma \times S$ is a 
%ternary relation,
%denoted as the set of {\em labeled transitions}. 
\end{definition}

%Note 
% that an (unlabeled) transition system can be seen as a labeled transition system where the set of labels consists of only one element.

We prefer to use infix notation and write $s \rightarrow s'$ to denote a {\em transition} from state $s$ to state $s'$ (i.e., $ (s,s') \in  {\rightarrow} $).
We write $\rightarrow^{k}$, $\rightarrow^{+}$, $\rightarrow^{=}$, $\rightarrow^{*}$
for the $k$-step iteration of $\rightarrow$, its transitive closure, its reflexive closure, its reflexive and transitive closure.
Let $X,Y \subseteq S$ and $k \in \mathbb{N}$. We denote $X \longrightarrow^{*} Y$ (resp. $X \longrightarrow^{\leq k} Y$) if from all states $x \in X$ there exists a path (resp. of length smaller than $k$) that reaches a state $y \in Y$.
\noindent
The set of {\em (immediate) successors} of a state $s \in S$ is deﬁned as $\post(s) = \{ s' \in S \mid  ~ s \xrightarrow{} s'\}$. A state without successors is called a {\em dead end}. The set of {\em (immediate) predecessors} of a state $s \in S$ is deﬁned as
$\pred(s) = \{ s' \in S \mid  ~ s' \xrightarrow{} s\}$.

%
\alain{enlever ce qui est inutile}
%
We use similar notation for $\post$ and $\pred$...
%
Determinism, (infinite) paths, their length, and concatenation in unlabeled transition systems are
then defined as expected.

@@

@@

@@

We abbreviate $x \leq y \not\leq x$ by $x < y$.
% A partial ordering (a po) is an antisymmetric qo
\alain{Any qo induces an equivalence relation ($x \equiv y$ iff $x \leq y \leq x$).}
% and gives rise to a po between the equivalence classes ?
% 	do we need po ?

@@

\begin{definition}
 A {\em well-quasi-ordering} (wqo) is any quasi-ordering $(X,\leq)$ such that, for any infinite sequence $x_0, x_1, x_2, ...$ in $X$, there exist indexes $i \leq j$ with $x_i \leq  x_j$.
\end{definition}
